---
title: "sugr"
subtitle: "Preparation of Continouos Glucose Measurements for usage in a shiny app"
author: "Tobias"
date: "`r format(Sys.time(), '%d. %b, %Y')`"
---

::: tldr
**tl;dr:** CGMs are a great source of (timeseries) data to tinker with. Here I describe the process of cleaning up and preparing the data of Medtronics systems in order to use it in a shiny app.
:::

```{r init}
#| include: false
library(tidyverse)
```

# Abstract

I'm a type I diabetic and therefore use an insulin pump and a connected CGM - continouos glucose measurement - system from Medtronic. The company kindly provides access to the raw data, even though only by the use of propriatory software and a web interface. This means I can't access the data through some sort of API to automate the collection thereof and have to run through that process manually on a regular basis.

Apart from that minor nuisance what's more important though is the fact that the data needs some touch-ups before you can actually use them:

- data from the pump and the GCM sensor are basically two different datasets but kept in the same csv
- while the time interval is 5 minutes (a reading from the GCM), not much else is going on otherwise
- direct blood glucose measurements, insulin doses, etc. are delivered _whenever_ and not on-point of the 5min intervals

Here I'll walk through the process of cleaning that up.

# Reading the Raw Data

First let's have a look at what we've got here. To save ourselves some headaches, I already put all the settings in all the places such that we get a nice tabular dataset. Most should be pretty self-explanatory: entries are delimited by a semicolon, the date format has to be sepcified and so on. I set `skip = 6` because, well, that's where the actual csv entries actually begin. Beforehand it's just some unstructured data about the device, patient, date of recordng, etc.

```{r b1_read_data}
data_raw <- read_delim("data/sugr/carelink-export-220217.csv", 
                  delim = ";", escape_double = FALSE, 
                  col_types = cols(Date = col_date(format = "%Y/%m/%d"), 
                                   Time = col_time(format = "%H:%M:%S")), 
                  locale = locale(decimal_mark = ","), 
                  trim_ws = TRUE, skip = 6)

head(data_raw)
```

Okay, so, a **Warning**. What's that about?

```{r b1_warning}
problems(data_raw) %>% select(-file) %>% print(n = 120)
```

Alright, the first ~70ish are actual parsing errors that lead to empty cells. Not great, but also no entries of huge importance^[mostly hard-coded things like ratios or target values that change slightly during a day but stay the same across all days], as far as I can tell. The rest though result from the circumstance that the file is basically seperated in CGM sensor data and insulin pump data, as mentioned in the beginning. Luckily though, `readr`/`vroom` just ignores the two rows where the new table begins - a device description and exactly same column names in the same order as before - and just continues parsing the CGM data. Let's take a look at those whopping 51 Variables:


```{r b1_glimpse_data}
glimpse(data_raw)
```

That's a whole lotta `<NA>`s! Apart from that, it looks pretty decent already. Here's what needs to be done still:

- throw out the variables we don't need (spoiler: most of them)
- clean up the column names
- `time` precision by minutes does absolutely suffice
- a combined `datetime` column would be nice
- a column of weekdays, too, just for some quality of life improvement

It also totally makes sense to split pump from CGM data, of course. Since both are structurally identical, we'll do that after the clean-up.

# The Clean-up Procedure

There are basically three groups of variables:

1. those of our main interest, like glucose measurements, insulin dosages, carbohydrate intake, etc
2. some that might be of minor interest, but rather when it comes to device usage and diagnostics than actual medical interest, like alarms, suspension and rewind times, etc
3. not interesting or useful at all

For now, we'll focus on the first type only.
